---
title: "vpb-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vpb-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(vpb)
```

VPB package is used to create the Vector Persistence Blocks for the gigen topology diagram. You can find more details on the topic in the article

    Kit C. Chan, Umar Islambekov, Alexey Luchinsky, Rebecca Sanders, "A computationally efficient framework for vector representation of persistence diagrams", arXiv:2109.08239

Let us first create the some sample data set:

```{r}
# sample 100 points uniformly from unit circle and add Gaussian noise
N <- 100
set.seed(123)
X <- TDA::circleUnif(N) + rnorm(2*N,mean = 0,sd = 0.2)
plot(X, asp=1)
```

We can use TDA package to create topology diagram from it:

```{r}
D <- TDA::ripsDiag(X,maxdimension = 1,maxscale = 2)$diagram
```

```{r}
# switch from birth-death to birth-persistence coordinates
D[,3] <- D[,3] - D[,2] 
colnames(D)[3] <- "Persistence"
head(D)
```

```{r}
library(ggplot2)
library(magrittr)
```

Here is the plot of this presistence diagram

```{r}
df <- data.frame(dim=as.factor(D[,1]), birth=D[,2], persistence = D[,3])
df %>% 
  ggplot(aes(x=birth, y=persistence, shape=dim, color = dim)) + geom_point() +
  geom_abline(intercept = 0, slope = 1)
```

Let us now compute the Vector Persistence Block of this diagram with homological dimension H0 and tau=0.3

```{r}
ySeqH0 <- unique(quantile(D[D[,1]==0,3],probs = seq(0,1,by=0.1)))
T0 <- computeVPB(D,homDim = 0,xSeq=NA,ySeqH0,tau = 0.3)
T0
```

The same result can be achieved using the C++ version of the function

```{r}
T0_C <- computeVPBcpp(D,homDim = 0,xSeq=NA,ySeqH0,tau = 0.3)
T0_C

```

Here it the H1 result

```{r}
xSeqH1 <- unique(quantile(D[D[,1]==1,2],probs = seq(0,1,by=0.2)))
ySeqH1 <- unique(quantile(D[D[,1]==1,3],probs = seq(0,1,by=0.2)))
computeVPB(D,homDim = 1,xSeqH1,ySeqH1,tau = 0.3) 
computeVPBcpp(D,homDim = 1,xSeqH1,ySeqH1,tau = 0.3) # compare with C++ version
```

